---
title: "penguin flipper lengths ANOVA"
author: "Adrienne Hewitt"
date: "November 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in the penguins.csv data

Packages: tidyverse, car, onewaytests

```{r}

library(tidyverse)
library(car)
library(onewaytests)

penguins<-read_csv("penguins.csv")

penguins

```


Question: Is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelie penguins near Palmer Archipelago?


Maybe I can use a one-way ANOVA...
-independent observations
-equal variance (Levene's test)
-normality (histograms, qqplots)

```{r}

counts<- with(penguins, table(species, sex))

penguins2<-penguins %>%
  mutate(
    common_name=
      case_when(
        species=="Chinstrap penguin (Pygoscelis antarctica)"~"Chinstrap",
        species=="Gentoo penguin (Pygoscelis papua)"~"Gentoo",
        species=="Adelie Penguin (Pygoscelis adeliae)"~"Adelie"
      )
  ) %>% 
  select(-species) %>% 
  filter(sex=="FEMALE")
```


Exploratory graphs and a Levene's test for equal variances
```{r}

hists<-ggplot(penguins2, aes(x=flipper_length))+
  geom_histogram(aes(fill=common_name))+
  facet_wrap(~common_name, scale="free")

hists

qqplots<-ggplot(penguins2, aes(sample=flipper_length))+
  geom_qq(aes(color=common_name))+
  facet_wrap(~common_name, scale="free")

qqplots

#based on qqplots and histograms, observations for each group appear normally distributed. Also, n>30 for all groups, so means will be normally distributed by CLT. 


```

Evaluate variances and perform a Levene's test for equal variances

```{r}

#If the largest sample variances is <4x greater than the smallest sample variance, then these are "close enough" to consider variances equal

variances<-penguins2 %>% 
  group_by(common_name) %>% 
  summarize(variance=var(flipper_length))

variances #the largest variance (33.1) is not more than 4x greater than the smallest variance (15.2)

#Levene's test
#leveneTest function in the 'car' package

levene_penguin<-leveneTest(flipper_length~common_name, data=penguins2)
levene_penguin

#H0: No difference in variances/variances are equal
#HA: Significant difference in variances/variances not equal

#There is no significant difference in variances across species

```


One-way ANOVA for penguin flipper length, by species (common_name)
```{r}

#H0: means across all groups are equal
#HA: at least two group means differ significantly

penguin_aov<-aov(flipper_length~common_name, data=penguins2)
summary(penguin_aov)

#Post-hoc testing by Tukey's HSD
#H0: No sig diff in pairwise means
#HA: Sig diff in pairwise means

penguin_ph<-TukeyHSD(penguin_aov)
penguin_ph

#aov.test function in the 'onewaytests' package does allow for in-line referencing

```


Mean flipper lengths of femail penguins differed significantly for all species studied (one-way anova, F(df)= statistic, p<0.001, $\alpha$=0.05 with post-hoc Tukey's HSD, $\alpha$=0.05)